---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rr4r: Rust regex for R

<!-- badges: start -->
[![R-CMD-check](https://github.com/yutannihilation/rr4r/workflows/R-CMD-check/badge.svg)](https://github.com/yutannihilation/rr4r/actions)
<!-- badges: end -->

rr4r is my practice to use Rust in an R package with the power of [extendr](https://github.com/extendr/extendr).
My current goal is to create the same interface to stringr's functions using Rust's [regex crate](https://docs.rs/regex/1.4.3/regex/).

## Installation

``` r
remotes::install_github("yutannihilation/rr4r")
```

## Example

### `rr4r_detect()`

```{r detect}
library(rr4r)

fruit <- c("apple", "banana", "pear", "pinapple", NA)

rr4r_detect(fruit, "a")
rr4r_detect(fruit, "^a")
rr4r_detect(fruit, "a$")
rr4r_detect(fruit, "b")
rr4r_detect(fruit, "[aeiou]")
```

### `rr4r_extract()`

```{r extract}
shopping_list <- c("apples x4", "bag of flour", "bag of sugar", "milk x2")

rr4r_extract(shopping_list, "\\d")
rr4r_extract(shopping_list, "[a-z]+")
rr4r_extract(shopping_list, "[a-z]{1,4}")
rr4r_extract(shopping_list, "\\b[a-z]{1,4}\\b")
```

`rr4r_extract()` can accept `i` to choose which capture group to extract.
This is useful when you don't need the whole match (with stringi/stringr,
you can use lookahead or lookbehind for this purpose, but Rust's regex crate
doesn't support it for performance reason).

```{r extract_index}
# Without index
rr4r_extract(c("<p>foo</p>", "<p>bar</p>"), "<p>(.*)</p>")

# With index
rr4r_extract(c("<p>foo</p>", "<p>bar</p>"), "<p>(.*)</p>", 1)

# stringr's equivalent using positive lookahead/lookbehind
stringr::str_extract(c("<p>foo</p>", "<p>bar</p>"), "(?<=<p>).*(?=</p>)")
```


### `rr4r_extract_all()`

```{r extract_all}
# Extract all matches
rr4r_extract_all(shopping_list, "[a-z]+")
rr4r_extract_all(shopping_list, "\\b[a-z]+\\b")
rr4r_extract_all(shopping_list, "\\d")
```


### `rr4r_extract_groups()`

```{r extract_groups}
library(dplyr, warn.conflicts = FALSE)

tibble(
  x = c("2020-01-02", "2021-12-31", "2022-08-09")
) %>% 
  mutate(
    rr4r_extract_groups(x, 
      pattern = "(?P<year>\\d{4})-(?P<month>\\d{2})-(?P<day>\\d{2})"
    )
  )
```


### `rr4r_replace()`

```{r replace}
fruits <- c("one apple", "two pears", "three bananas")
rr4r_replace(fruits, "[aeiou]", "-")

rr4r_replace("a1", "\\D", toupper)
```

```{r replace2}
library(lubridate)

x <- c("Today is 2020-02-02", "2021-03-30")
ptn <- "(?P<year>\\d{4})-(?P<month>\\d{2})-(?P<day>\\d{2})"

one_month_ago <- function(year, month, day) {
  d <- make_date(year, month, day) %m-% months(1)
  format(d, "%Y-%m-%d")
}

rr4r_replace(x, ptn, one_month_ago)
```


